---
title: "Discussion_Week9_MQari"
author: "Mubashira Qari"
date: "2024-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Chapter 9

### Question 2:

Let S200 be the number of heads that turn up in 200 tosses of a fair coin.

Estimate

(a) P(S200 = 100).
(b) P(S200 = 90).
(c) P(S200 = 80).

### Solution:

Using the Binomial Probability function


\[ P(x=k)= \left( \begin{bmatrix}
    n \\
    k
\end{bmatrix}\right)p^k(1-p)^{n-k}\]

The binomial distribution model is used for finding the probability of success of an event which has only two possible outcomes in a series of experiments.

R has four in-built functions to generate binomial distribution. 

### Solution (a):

```{r}
# number of trial = 200

# probability of head = 0.5 

# number of head = 100 

n <- 200
p <- 0.5
k <- 100 

dbinom<-dbinom(k,n,p)
dbinom
```
### Solution (b):

```{r}
# number of trial = 200

# probability of head = 0.5 

# number of head = 100 

n <- 200
p <- 0.5
k <- 90 

dbinom<-dbinom(k,n,p)
dbinom
```
### Solution (c):

```{r}
# number of trial = 200

# probability of head = 0.5 

# number of head = 100 

n <- 200
p <- 0.5
k <- 80 

dbinom<-dbinom(k,n,p)
dbinom
```

### Page 338 

### Question 3:

A true-false examination has 48 questions. June has probability 3/4 of answering a question correctly. April just guesses on each question. A passing score is 30 or more correct answers. Compare the probability that June passes the exam with the probability that April passes it.

### Solution:

pbinom(x, size, prob) - the cumulative probability of an event. It is a single value representing the probability.

x is a vector of numbers.

p is a vector of probabilities.

n is number of observations.

size is the number of trials.

prob is the probability of success of each trial.

Mean = n * p
Variance = n * p * (1 - p)


Calculating the probability that June passes the exam

```{r}
1 - pnorm(30.5, mean=48*3/4, sd=sqrt(48*3/4*1/4))
```

Calculating the probability that April passes the exam

```{r}
1 - pnorm(30.5, mean=48*1/2, sd=sqrt(48*1/2*1/2))
```

