---
title: "Data 605 Discussion Week1"
author: "Mubashira Qari"
date: "2024-01-28"
output: html_document
---

### Unitary Matrices: Solving system of Equation with Inverse Matrix

Compute the inverse of the coefficient matrix of the system of equations below and use the inverse to solve the system.

4x + 10y = 12
2x + 6y = 4

The inverse of a matrix plays the same roles in matrix algebra as the reciprocal of a number and division does in ordinary arithmetic: 

we can solve a matrix equation like Ax=b
for the vector x by multiplying both sides by the inverse of the matrix A

Ax=b ⇒ A−1Ax=A−1b ⇒ x=A−1b

The following system of equation illustrate the basic properties of the inverse of a matrix.


### Load the matlib package

```{r}
library(matlib)

```
### Creating a 2 x 2 matrix

```{r}
A <- matrix(c(4, 2, 10, 6), 2, 2)
b <- c(12,4)
showEqn(A, b)
```
# Finding Rank

```{r}
A

c( R(A), R(cbind(A,b)) )    

```
```{r}
# Finding the determinant of A
det(A)
```
# Finding the inverse of A

```{r}

AInv <- inv(A)

AInv

```
```{r}
# Proving A inverse X A is Identity matrix

 AInv %*% A


```


### Identity Matrix * X is X

```{r}
# I * X = AInv * b => X = AInv * b

X = AInv %*% b

X
```
# Checking the answer

```{r}
Solve(A, b, fractions = TRUE)
```
### Plot the equations

```{r}

plotEqn(A,b)

```
### This intersection of two lines proves that this consistant system of equation has at least one solution.

