---
title: "Discussion_Week_6_MQari"
author: "Mubashira Qari"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Chapter3, Page 97

### Question 12: A symphony orchestra has in its repertoire 30 Haydn symphonies, 15 modernworks, and 9 Beethoven symphonies. Its program always consists of a Haydnsymphony followed by a modern work, and then a Beethoven symphony.

(a) How many different programs can it play?

(b) How many different programs are there if the three pieces can be played
in any order?

(c) How many different three-piece programs are there if more than one
piece from the same category can be played and they can be played in
any order?


### (a) How many different programs can it play?

```{r}
# Following are the three symphonies
Haydn <- 30
modernworks <- 15
Beethoven <- 9
```


```{r}
total_prog <- Haydn*modernworks*Beethoven

total_prog
```
### (b) How many different programs are there if the three pieces can be played in any order?

Following combinations can be possible:

Set1:Haydn,Modern,Beethoven
Set2:Haydn,Beethoven,Modern
Set3:Beethoven,Modern,Haydn
Set4:Beethoven,Haydn,Modern
Set5:Modern,Beethoven,Haydn
Set6:Modern,Haydn,Beethoven

Since there are 6 possible way

```{r}
possible_prog <- 6 * total_prog

possible_prog <- 6 * 4050

possible_prog

```
### (c) How many different three-piece programs are there if more than one piece from the same category can be played and they can be played in any order?



```{r}
sum <- Haydn + modernworks + Beethoven

total_ways<- factorial(sum)/factorial(sum-3)

total_ways
```

