---
title: "Discussion_Week8_MQari"
author: "Mubashira Qari"
date: "2024-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 3 Let X1 and X2 be independent random variables with common distribution
pX =

0 1 2
1/8 3/8 1/2


\[P_x
 = \left(
    0 \;1 \;2 \\
    3/8 \;1/2 \;1/8  
 \right) \]


### Solution:

Using the convolution of the above distribution to find the probability distribution:

When Z=0, both \[X_{1} = 0\] and \[X_{2}=0\]

\[ P(Z=0)= P_x(0)*P_x(0) = \frac{1}{8} * \frac{1}{8}= \frac{1}{64}\]

```{r}
p0 <- 1/8 * 1/8

p0 == 1/64
```
When Z=1, \[X_{1} = 0\] and \[X_{2}=1\]

\[ P(Z=1)= m(0)*m(1) + m(1)*m(0) = \frac{1}{8} * \frac{3}{8} + \frac{3}{8} * \frac{1}{8}= \frac{6}{64}\]

```{r}
p1 <- (1/8 * 3/8) + (3/8 * 1/8)

6/64 == p1
```
When Z=2,

\[ P(Z=2)= m(0)*m(2) + m(1)*m(1) + m(2)*m(0)= \frac{1}{8} * \frac{1}{2} + \frac{3}{8} * \frac{3}{8} +\frac{1}{2} * \frac{1}{8}= \frac{17}{64}\]

```{r}
p2 <- (1/8 * 1/2) + (3/8 * 3/8) + (1/2 * 1/8)

17/64 == p2
```
When Z=3, 

\[ P(Z=3)= m(1)*m(2) + m(2)*m(1) = \frac{3}{8} * \frac{1}{2} + \frac{1}{2} * \frac{3}{8}= \frac{24}{64}\]

```{r}
p3 <- (3/8 * 1/2) + (1/2 * 3/8)

24/64 == p3
```
When Z=4, both  \[X_{1} = 2\] and \[X_{2}=2\]

\[ P(Z=4)= \frac{1}{2} * \frac{1}{2}= \frac{1}{4}= \frac{16}{64}\]

```{r}
p4 <- (1/2 * 1/2)

16/64 == p4
```
Probability distribution (sum of each discrete distribution value above) sums to 1:

```{r}
prob_sum <- p0 + p1 + p2 + p3 + p4

1.0 == prob_sum
```

