---
title: "HW10_MQari"
author: "Mubashira Qari"
date: "2024-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question:

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy).

(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).

(c) Which strategy gives Smith the better chance of getting out of jail?

### Solution:

###(a) Calculating Probability of getting $8 with Timid Strategy (he bets 1 dollar each time):

Applying the the probability of reaching the winning sum of $8 from the gambler’s ruin section of our textbook (Chapter 12.2):

\[P = \frac{(\frac{q}{p})^S-1}{(\frac{q}{p})^M-1}\]

S - Smith's starting bet
q - is the probability of failure in a given turn
p - is the probability of success in a given turn

```{r}
p <- 0.4
q <- 0.6
M <- 8
S <- 1

(p_timid <- ((q / p)**S - 1) / ((q / p)**M - 1))

```
Loading Markov Chain Library 

```{r}
library(markovchain)
```
### Applying the Bold Strategy (Each time he bets as much as possible)

In this case the amount gambled depends on the amount of money Smith has in his possession 

First, he starts with s=$1 (k=1)

Next, if he wins, he will have $2

Below are the the possible winning amounts he can have at any point in the game are $0,$1,$2,$4,$8
.
.
```{r}
transition_matrix <- matrix(c(1, 0, 0, 0, 0,
                              0.6, 0, 0.4, 0, 0,
                              0.6, 0, 0, 0.4, 0,
                              0.6, 0, 0, 0, 0.4,
                              0, 0, 0, 0, 1), ncol=5,nrow=5)

rownames(transition_matrix) <- c("$0", "$1", "$2", "$4", "$8")
colnames(transition_matrix) <- c("$0", "$1", "$2", "$4", "$8")

#  only has an one entry in the $1 position

starting_state <- c(0, 1, 0, 0, 0)
```

To win the game, Smith needs to win 3 times in a row.

If he loses at any point, he goes bust (because under the bold strategy, he is betting his sum until $4, after which he either wins or loses).

$\1→$2→$4→$8

The probability of Smith ending with $8 ends up at 0.064
.

```{r}
x <- starting_state
for (i in 1:3){
  x <- transition_matrix %*% x
 
}

(p_bold <- unname(x["$8", ]) )
```
### (c) Strategy that gives Smith the better chance of getting out of jail:

Higher probability of Smith reaching his target value than the timid strategy.

```{r}
p_bold > p_timid
```

