---
title: "Data 605 HW2"
author: "Mubashira Qari"
date: "2024-01-29"
output:
  html_document: default
  pdf_document: default
---

## PROBLEM SET 1:

### (1) Show that ATA != AAT in general. (Proof and demonstration.)

### Load the Packages

```{r}
library(matrixcalc)
library(matlib)

```
### Creating a 3 X 3 Matrix

```{r}
A = matrix(c(1, 1, 2, 3, 2, 2, 1, 1, 1), nrow = 3)
A
```

### Finding the determinant of A to check: if det(A) != 0 then inverse exists

```{r}

det(A)

```
### Finding the tranpose of A

```{r}
AT <- t(A)

AT

```
### Mutiplying A AI

```{r}
A %*% AT
```
### Mutiplying AI A

```{r}

AT %*% A

```
### Checking for proof and it proves that (A X AI != AI X A)

```{r}
(A %*% AT == AT %*% A)

```
### (2) For a special type of square matrix A, we get AT A = AAT. Under what conditionscould this be true? (Hint: The Identity matrix I is an example of such a matrix).


### Creating an identity matrix for checking the case of a special square matrix

```{r}
I <- diag(3)

I
```
### Cretaing transpose of A

```{r}
A <- I
AT <- t(A)
AT
```


```{r}
AT %*% A 

```


```{r}
A %*% AT

```
For a special type of square matrix called Identity Matrix, we get AT A = A AT

```{r}

(A %*% AT == AT %*% A)

```

## PROBLEM SET 2:

```{r}
A = matrix(c(1, 4, 3, 1,3, 5, 1, -1, 3), nrow = 3)
A

```
### In order to eliminate the 2 in the 2nd row, we can set up the following elimination matrix E21:

```{r}
E21 = matrix(c(1,-4,0,0,1,0,0,0,1),nrow=3)
E21
```


```{r}

E21 %*% A


```
### To eliminate 3 in row 3, we can choose E32 to be:


```{r}
E31 = matrix(c(1,0,-3,0,1,0,0,0,1),nrow=3)
E31
```


```{r}

E31 %*% E21 %*% A


```
### Notice that the pivot in row two, a22 = −1. This leads us to choose the following elimination matrix E32:

```{r}
E32 = matrix(c(1,0,0,0,1,2,0,0,1),nrow=3)
E32
```
### This gives us the U matrix

```{r}
U <- E32 %*% E31 %*% E21 %*% A
U
```
### We can compute L as the inverse of the product of the elimination matrices.

### L = INV(E32 × E31 × E21)  

```{r}

inv(E32)
inv(E31)
inv(E21)
```
### We can also use the solve function in R to get the inverse of the matrices

```{r}
L <- solve(E21) %*% solve(E31) %*% solve(E32)

L
```
### Finally, we will solve for proof that A = LU

```{r}
L %*% U
```
### This proves that A = LU

```{r}
(L %*% U == A)

```




