---
title: "Untitled"
author: "Mubashira Qari"
date: "2024-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### With the attached data file, build and visualize  eigenimagery that accounts for 80% of the variability.  Provide full R code and discussion.
 

```{r}
library(doParallel)
library(foreach)
library(iterators)
library(jpeg) #for readJPEG()
library(EBImage) #for the resize() function
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")

```


```{r}
	
BiocManager::install("EBImage")

```


```{r}
files=list.files(path='C:/Users/Uzma/OneDrive/Desktop/Documents/jpg',pattern="\\.jpg")
files
```
### Rescaling the iamges for viewing by initializing a function

```{r}
height=1200
width=2500
scale=20
plot_Image = function(path, add=FALSE) #initialize function
{
  require('jpeg')
  jpg = readJPEG(path, native=T) # read the file
  res = dim(jpg)[2:1] # get the resolution, [x is 2, y is 1]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]), #set the X Limits by size
         asp=1, #aspect ratio
         type='n', #don't plot
         xaxs='i',yaxs='i',#prevents expanding axis windows +6% as normal
         xaxt='n',yaxt='n',xlab='',ylab='', # no axes or labels
         bty='n') # no box around graph
  rasterImage(jpg,1,1,res[1],res[2]) #image, xleft,ybottom,xright,ytop
}
```

### Loading the images data into an array

```{r}
#initialize array with zeros.
shoeIMG=array(rep(0,length(files)*height/scale*width/scale*3),
         #set dimension to N, x, y, 3 colors, 4D array)
         dim=c(length(files), height/scale, width/scale,3)) 

for (i in 1:length(files)){
  #define file to be read
  tmp=paste0("C:/Users/Uzma/OneDrive/Desktop/Documents/jpg/", files[i])
  #read the file
  temp=EBImage::resize(readJPEG(tmp),height/scale, width/scale)
  #assign to the array
  shoeIMG[i,,,]=array(temp,dim=c(1, height/scale, width/scale,3))
}

```

### Plotting the images of shoes which are in 3 X 3 table

```{r}
par(mfrow=c(3,3)) #set graphics to 3 x 3 table
par(mai=c(.3,.3,.3,.3)) #set margins 
for (i in 1:8){  #plot the first images only
plot_Image(writeJPEG(shoeIMG[i,,,]))
}

```
### Principal Component Analysis:

### Standardizing the data by redimensioning it into two dimensional matrix

```{r}
height=1200
width=2500
scale=20

scaled_Data=shoeIMG

dim(scaled_Data)=c(length(files),height*width*3/scale^2)
dim(scaled_Data)
```
### Calculating the Principal Components by setting the Scores and Corelation TRUE 

### where the scores represent the eigen vector

```{r}
pca_Result = princomp(t(as.matrix(scaled_Data)), scores=TRUE, cor=TRUE)

pca_Result
```

### Varifying that the sum of Varience = 1 


```{r}
variance = sum(pca_Result$sdev^2/sum(pca_Result$sdev^2))

variance
```

### Finding the nnumber of components that accounts for 80% of variability 

### First 5 component accounts for 80% of variability

```{r}
final_Comp = pca_Result$sdev^2/sum(pca_Result$sdev^2)

sum(final_Comp[1:3])
```

### First 10 component accounts for 90% of variability

```{r}
sum(final_Comp[1:6])
```
### Tranposing the Eigen Vector Scores to repoduce matrix for replotting

```{r}
pca_result2=t(pca_Result$scores)
dim(pca_result2)=c(length(files),height/scale,width/scale,3)
par(mfrow=c(5,5))
par(mai=c(.001,.001,.001,.001))
for (i in 1:17){  #plot the first 81 Eigenshoes only
plot_Image(writeJPEG(pca_result2[i,,,], quality=1,bg="white"))
}
```
### These are the most important images where  the first image accounts for the most variability among others.
