---
title: "HW7_MQari"
author: "Mubashira Qari"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 1. Let X1, X2, . . . , Xn be n mutually independent random variables, each ofwhich is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

### Solution:

Given: X1, X2, . . . , Xn be n mutually independent random variables.

Each of which are uniformly distributed on the integers from 1 to k.

Y denote the minimum of the Xi’s.

To find the distribution of Y, we will find the probability that Y takes on each possible value from 1 to k, 

where y is the range of all possible value from 1 to k.

For that we will' calculate the  cumulative distribution functions (CDFs) and probability mass functions (PMFs).

Let PMF of Y is P(Y=y) where y is the range of all possible value from 1 to k.

The probability that P(Y>y) is the probability that all Xi's are greater than y.

Also, since Xi's are are independent and uniformly distributed from 1 to k, the probability that any single Xi is greater than y is:

\[
 = \frac{k-y}{y}\; \].

The probability of all Xi's greater than y is:

\[
 \left( \frac{k-y}{y}\; \right)^n\].

The CDF of Y:

where CDF is a function that describes the probability that a random variable (continuous or discrete) will take on a value less than or equal to a certain value. 

\[F_{Y}(y) =  P(Y \leq y) = 1 - P(Y > y) = 1 - \left( \frac{k-y}{k}\; \right)^n \]


Now to find the PMF of Y:

where PMF is a function that describes the probability of a discrete random variable taking on a certain value.

\[P(Y = y) = F_{Y}(y) - F_{Y}(y-1) \]

\[P(Y = y) = \left(1 - \left( \frac{k-y}{k}\; \right)^n \right) -  \left(1 - \left( \frac{k-(y-1)}{k}\; \right)^n \right)\]


\[P(Y = y) = \left(\frac{k-(y-1)}{k}\; \right)^n  -  \left(\frac{k-y}{k}\; \right)^n \]

The PMF of Y where Xi's is the minimum value and uniformly distributed from 1 to K. 

\[P(Y = y) = \frac{(k-y+1)^n -(k-y)^n}{k^n}\; \]



### 2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

### Solution:

To model the time until the occurrence of a certain event, such as machine failure, we will use the exponential distribution.


The probability density function (PDF) of the exponential distribution is:

\[f(x) = \lambda * e^{-\lambda*x}\; \]

where λ is the rate parameter (1/mean) 

x is the time until failure.


```{r}
lambda <- 1/10
```

\[X \sim Geom(\frac{1}{10})\]


\[P(Y > n) = 1 - P(X \leq 8) = 1 - \sum_{i=1}^{8}(0.9)^{i-1}(0.1) \]

\[1 - 0.1(1+0.9+0.9^2+...+0.9^7)\]

\[E[X]= \frac{1}{\frac{1}{10}}= 10 \]

\[\sqrt{a}=b \]

\[\sigma = \sqrt{Var(X)}= \sqrt{90} = 3 \sqrt{10}\]


a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)


Probability that the machine will fail after 8 is 43%,

```{r}
prob <- 1/10

year=8

round((1 - pgeom(year-1, prob)),4)

```
The Standard Deviation is 9.49.

```{r}
std_dev <- round((sqrt((1-prob)/prob^2)),2)
std_dev



```
b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.

Probability that the machine will fail after 8 is 45%,

```{r}
x <- 8
prob <- 1/10
round((1 - pexp(x, prob)), 4)
```
Standard deviation is 10.

```{r}
sqrt(1/prob^2)
```

c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years) 

Probability that the machine will fail after 8 is 43%, 

```{r}
x <- 0
year <- 8
prob <- 1/10
round((dbinom(x, year, prob )), 2)
```

The expected value is 0.80

```{r}
expected_value <- year*prob
expected_value
```
Standard deviation is 0.849.

```{r}
std_dev <- sqrt(year*prob*(1-prob))
round(std_dev,3)
```
d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

There is a 44.9% failure rate

```{r}
x <- 0
year <- 8
prob <- 1/10
t <- 1 
round((dpois(x, year*prob/t)), 3)
```
Expected Value is 0.8

```{r}
exp_val <- year*prob/t

exp_val
```
Standard Deviation is:

```{r}
(year*prob/t)^.5
```
